---
title: "Plotting"
author: "Erle Holgersen"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

For visualization, I focus on writing functions for plots I use frequently and that require many lines of base R. At present, the package support two plots, `km.plot` and `stripboxplot`.

# Kaplan-Meier Plots

Kaplan-Meier plots are used for visualizing censored data. They are most commonly used in medical studies to visualize differences in patient survival. 

```{r}
library(hedgehog);
library(survival);

# using example from survival package
km.plot(
    Surv(futime, death) ~ sex, 
    mgus
    );
```

```{r}
km.plot(
    Surv(futime, death) ~ sex, 
    mgus,
    show.risk.table = TRUE
    );
```

# Box plots

Standard box plots hide the sample size in each group. One way of getting around this is to add a background stripplot that shows each point in the dataset. 

```{r}
test.data <- data.frame(
    x = sample(letters[1:3], 100, replace = TRUE),
    y = rnorm(100)
    );

stripboxplot(y ~ x, test.data);
```
